### 2019-12-05 paper/paper.tex

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}

\title{A Systematic Approach to Analyzing Syndication of Viral News Stories}

\author{
 Gaurav Deshpande \\
  Institute for Software Research\\
  Carnegie Mellon University\\
  Pittsburgh, PA 15213 \\
  \texttt{gdeshpan@andrew.cmu.edu} \\
  %% examples of more authors
 \And
    Sam E. Teplov\\
    %\thanks{Use footnote for providing further
    %information about author (webpage, alternative
    %address)---\emph{not} for acknowledging funding agencies.} \\
  Institute for Software Research\\
  Carnegie Mellon University\\
  Pittsburgh, PA 15213 \\
  \texttt{steplov@andrew.cmu.edu} \\
 \And
   Alon Peer \\
  Department of Computer Science\\
  Carnegie Mellon University\\
  Pittsburgh, PA 15213 \\
  \texttt{apeer@andrew.cmu.edu} \\
 \AND
    Dr. Nicolas Christin \\
    Department of Computer Science \\
    Carnegie Mellon University\\
    Pittsburgh, PA 15213 \\
   \texttt{nicolasc@andrew.cmu.edu} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
\maketitle

\begin{abstract}
In this work, we present a novel, semiautonomous methodology for analyzing syndicated news articles. We first detail our manual process of news syndication analysis from which we observe a number of unsettling patterns. From there, we go on to categorize and define two different forms of syndication: top-down and bottom-up. We then transform our manual process into a semiautomated one which combines web scraping and machine learning techniques to create and then analyze a dataset of viral news stories. From the analysis, we are able to make some inferences about the nature of syndicated news articles, as well as trace back a viral news story to its original publisher and article. By presenting this work, we hope to minimize the effects that blind news syndication has on the spread of misinformation.

\end{abstract}

% keywords can be removed
\keywords{Fake news \and News propagation \and News syndication}

\section{Introduction}

News are everywhere- television, newspapers, social media. We consume it whether we like it or not. But the question is, how do we know which news are true, and which are false?

Fake news has been a real issue since the spark of social media, especially in the recent years, with its spike coming during the 2016 U.S. presidential election. In fact, a study from Ohio State University shows that fake news had a significant impact on the results of the elections. The researchers found a correlation between the amount of Obama voters who believed fake news about Clinton and the amount of them who chose to vote for Trump. At that time, the media was swarmed with fake news, and one could hardly blame people who couldn't distinguish them from real news.

But what exactly is fake news? According to Wikipedia, "fake news is a form of news consisting of deliberate disinformation or hoaxes spread via traditional news media (print and broadcast) or online social media". In addition to this definition, fake news may also include clickbait stories, propaganda, satire, biased news, or even sometimes just bad journalism.

These days, anyone can publish content on the internet, whether it is on a website, a blog, or a social media profile, and potentially reach large audiences. The fact that people nowadays rely on the internet to consume news doesn't make this problem any easier to solve.
Fake news is very profitable for the publisher, as viral stories generate large amounts of traffic, thus increasing advertisement revenue. For that reason, fake news are widely common around the internet. 
Another motive to publish fake news is to promote a political agenda. Fake political news stories are meant to persuade consumers to accept a biased of false beliefs (like the fake news regarding Hillary Clinton during the 2016 U.S. presidential election). This is the main reason detecting fake news is such a challenge. These false stories were created to sound true and interesting.

A different variation of fake news is syndication. A website can take a real news story, and change it a bit, so it will fit that website's beliefs. A news story can be syndicated in various ways- changing the title, adding interpretations that aren't necessarily true, and much more. A popular form of syndication is news stories that are posted by local newspapers and websites, and as they get viral, these stories get picked up by bigger and more popular news websites. On the way from the local news source to the national (or international) news website, the story can get syndicated. Sometimes, it's not necessarily on purpose, but it happens due to the lack of due diligence by the mainstream media.

One can try and detect fake news by checking the source of the story, read beyond the title (which is intended to attract users and thus contains more interesting and controversial topics), try to trace the article back to the original source, or even make sure the article is not a satire by any chance.
Even though it's possible, doing so is not an easy task, and that is why there has been a lot of work put into automating this process. A step towards this, is detecting and analyzing the syndication of viral news stories. That is, how stories change through time and the difference between articles regrading the same stories but posted in different websites.

\section{Background}
The field of misinformation and fake news has become widely researched by academics ever since the Russian misinformation campaign against the 2016 United States presidential election came to light\cite{stelzenmuller2017impact}. The most straightforward approach to detecting fake news is applying various machine learning techniques, and this is what many researchers have focused on thus far.

\subsection{Fake News Detection}Shu et al. goes through a comprehensive overview of all of the various ways of attacking the problem of fake news detection and classification from a data mining perspective \cite{shu2017fake}. From the perspective of detecting fake news via machine learning techniques, Shu et al. outlines that there are generally four different types of features that can be used to detect fake news: news content features, linguistic-based features, visual-based features, and social context features. These features are then applied to one of two different types of models: news content models, or social context models . Just as they sound, news content models focus on fact checking claims in the news article or looking at stylistic features that may indicate that the article is fake. Social context models, on the other hand, focus on how users interact with the article, the virality of the article, and the propagation path of the news story. These two different methods are rarely used independently, but rather compliment each other in many hybrid approaches, such as ensemble models \cite{shu2017fake}.

\subsection{Open Issues in Fake News Detection}
Shu et al. closes by giving an overview of many of the various open areas of research in this field, as well as proposing some future research ideas. Shu et al. outlines four future research directions: data-oriented, feature-oriented, model-oriented, and application-oriented. Within the area of data-oriented research, Shu et al. brings forward a couple of issues that need to be addressed. For one, there is no large, encompassing fake news data set which can be used as a benchmark to facilitate future research \cite{shu2017fake}. In fact, this was such a large problem that in 2018, Shu et al. published another work detailing their methodology for building a large scale, open source, data set consisting of fake news, which is now available on Github \cite{shu2018fakenewsnet}. Another idea that Shu et al. proposes for future research within the data-oriented direction is early fake news detection \cite{shu2017fake}. Many of the current state-of-the-art fake news detection methods rely on data that needs to be accumulated over time and information that is not readily available when news is just starting to go viral \cite{liu2018early}. Shu et al. suggests creating a system that provides fake news alerts during the dissemination process so that misinformation can be flagged before it is viewed by a large audience. Another suggestion that is made is to look at social media posts made within some time delay of the original posts as a means of verifying the veracity of the post \cite{shu2017fake}.  
%\label{sec:headings}

\section{Related Work}
There has been work done on early fake news detection by analyzing propagation paths and time series. Most of this work, however, has been limited in scope in terms of the domains that it covers, as well as the type of media being analyzed. For example, past research has focused on a either a few social media platforms, or a specific media outlet \cite{liu2018early},\cite{jin2014news}. The research that has spanned across multiple different domains of news (social media, various news outlets, etc.) has been limited in scope when considering the type of media that is being analyzed \cite{zannettou2018origins}.

\subsection{Domain-Specific Approaches}

\subsection{Media-Specific Approaches}

\section{Methodology}
\subsection{Phase 1 - Manual Inspection}

To begin the process of analyzing syndication, we started off by manually gathering articles and seeing the propagation path of those articles. We tried to look for a pattern in the way articles change throughout time. This phase included x steps.

\subsubsection{Article Search}
To begin the manual inspection. We started looking through the internet for articles that gained a lot of traffic relative to other articles in the same website. We looked for a high number of shares or comments. Some of the news sites we looked at are "Daily Mail", "The Washington Post" and "Daily Caller".
\subsubsection{Original Source Search}
After we filtered out some articles due to content and popularity, we tried tracing the articles back to their original source. We google key words from the article, and that way we essentially got a list of different websites' articles of that news story. We paid close attention to the dates, to try and find the earliest article regrading said news story. 
\subsubsection{Propagation path}
Once we found the original news source, we looked at how the article title and content change over time and throughout different news sources. We tried to see how the website's beliefs changed the way the story was described in each and every news site, that is, how news got syndicated.

\section{Analysis}

\section{Contributions}

\section{Findings}

\section{Future Work}

\section{Conclusion}

\newpage

\bibliographystyle{unsrt}  
\bibliography{paper}  %%% Remove comment to use the external .bib file (using bibtex).
%%% and comment out the ``thebibliography'' section.

%%% Comment out this section when you \bibliography{references} is enabled.

\end{document}